Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-11-30 paper/sec-intro.tex
More work on draft
more work on paper

As in the case of single Higgs production,
in the SM the dominant mechanism for the production of a pair of Higgs bosons at the LHC is 
gluon fusion (see Ref.~\cite{baglio} and references therein), analogously to single Higgs production~\cite{Dittmaier:2012vm}.
%
For a center-of-mass energy $\sqrt{s} = 14\,$TeV, the next-to-next-to-leading order (NNLO)
total cross section is approximately $40\,$fb~\cite{deFlorian:2013jea},
which is increased by a further few percent once NNLL corrections
are accounted for~\cite{deFlorian:2015moa}.
%
Feasibility studies in the case of a SM-like Higgs boson
in gluon fusion have been performed for several different final states, including
$b\bar b\gamma\gamma$~\cite{Baur:2003gp,Barger:2013jfa},
$b\bar{b}\tau^+\tau^-$~\cite{Baur:2003gpa,Barr:2013tda,Dolan:2012rv,Dolan:2013rja},
$b\bar{b}W^+W^-$~\cite{Dolan:2012rv,Papaefstathiou:2012qe} and
$b\bar{b}b\bar{b}$~\cite{Baur:2003gpa,Dolan:2012rv,Wardrope:2014kya,deLima:2014dta}.
%
While these studies differ in their qualitative conclusions,
a consistent picture emerges 
that the ultimate precision in the determination of the Higgs trilinear
coupling $\lambda$
will only be achieved from the combination of various final states.

@@

The structure of this paper is the following.
%
In Sect.~\ref{mcgeneration} we discuss the modeling of the signal
and background processes using Monte Carlo event generators.
%
In Sect.~\ref{sec:analysis}
we introduce our analysis strategy, and in particular
the classification of each individual event into
different categories according to their topology.
%
The results for the traditional cut-based analysis
are presented in Sect.~\ref{sec:results}, and in
Sect.~\ref{sec:mva} we show how we can enhance signal
significance using multivariate techniques.
%
The robustness of our results as function of our assumptions
on the $b$-tag rate, mistag probability and momentum
resolution in quantified in Ref.~\ref{sec:optimisation}.
%
Finally in Sect.~\ref{sec:conclusions} we conclude and outline
how our results can be extended to estimate the accuracy
in the determination of $\lambda$ and the constraints on
BSM scenarios in an EFT framework.

>>>>

As in the case of single Higgs production,
in the SM the dominant mechanism for the production of a pair of Higgs bosons at the LHC is 
gluon fusion (see Ref.~\cite{baglio} and references therein), analogously to single Higgs production~\cite{Dittmaier:2012vm}.
%
For a center-of-mass energy $\sqrt{s} = 14\,$TeV, the next-to-next-to-leading order (NNLO)
total cross section is approximately $40\,$fb~\cite{deFlorian:2013jea},
which is increased by a further few percent once NNLL corrections
are accounted for~\cite{deFlorian:2015moa}.
%
Feasibility studies in the case of a SM-like Higgs boson
in gluon fusion have been performed for several different final states, including
$b\bar b\gamma\gamma$~\cite{Baur:2003gp,Barger:2013jfa},
$b\bar{b}\tau^+\tau^-$~\cite{Baur:2003gpa,Barr:2013tda,Dolan:2012rv,Dolan:2013rja},
$b\bar{b}W^+W^-$~\cite{Dolan:2012rv,Papaefstathiou:2012qe} and
$b\bar{b}b\bar{b}$~\cite{Baur:2003gpa,Dolan:2012rv,Wardrope:2014kya,deLima:2014dta,Barger:2013jfa}.
%
While these studies differ in their qualitative conclusions,
a consistent picture emerges 
that the ultimate precision in the determination of the Higgs trilinear
coupling $\lambda$
will only be achieved from the combination of various final states.

@@

The structure of this paper is the following.
%
In Sect.~\ref{mcgeneration} we discuss the modeling of the signal
and background processes using Monte Carlo event generators.
%
In Sect.~\ref{sec:analysis}
we introduce our analysis strategy, and in particular
the classification of each individual event into
different categories according to their topology.
%
The results for the traditional cut-based analysis
are presented in Sect.~\ref{sec:results}, and in
Sect.~\ref{sec:mva} we show how we can enhance signal
significance using multivariate techniques.
%
The robustness of our results under realistic pile-up
conditions expected at the HL-LHC is illustrated in
 Sect.~\ref{sec:pileup}.
%
Finally in Sect.~\ref{sec:conclusions} we conclude and outline
how our results can be extended to estimate the accuracy
in the determination of $\lambda$ and the constraints on
BSM scenarios in an EFT framework.

### 2015-11-30 paper/sec-mva.tex
more work on paper

\section{Results from the multivariate analysis}
\label{sec:mva}

From the cut-based analysis,
combining the significance of the three categories,
we end up with a signal significance of $S/\sqrt{B}\simeq 1.2(2.3)$
with all backgrounds (only QCD $4b$) taken into account.
%
This does not seem
to be enough to be able to claim observation
of Higgs pair production
in this channel, and even less to extract the trilinear Higgs  coupling
$\lambda$ with good precision.
%
In this section we show how the signal significance for
our process can
be substantially improved when
applying a multivariate analysis (MVA) to the events
 that survive the cut-based
 analysis.

%
Multivariate techniques are by now a common tool to enhance signal
significance in HEP analysis, and this is fully justified
since they open a new window to improve the performance
of many physics analysis and searches.
%
In particular, the classification of events as arising from either signal or
background processes via the use of a multivariate analysis is becoming
a mature
technique in LHC
applications~\cite{Baldi:2014pta,Aaltonen:2012qt,
  Wardrope:2014kya,Chatrchyan:2013zna,Dall'Osso:2015aia}.
%
For instance, the CMS $Vh(\to b\bar{b})$ analysis~\cite{Chatrchyan:2013zna}
is based on a Boosted Decision Tree (BDT) MVA, and the UCL $hh\to 4b$
feasibility study~\cite{Wardrope:2014kya}
also used  a BDT to improve signal discrimination.

In this section, first of all we present the specific MVA that we use,
based on feed-forward multi-layer neural networks.
%
We then introduce the input variables that are
used the MVA, including the jet substructure
variables, and then present the results of $S/\sqrt{B}$ including
the MVA effects.
%
Since we find that jet substructure variables are a crucial input
to the MVA, we end this section showing various kinematic distributions
of these variables.

@@

%
In our work we use a specific type of  MVA to
disentangle signal and background events in the $hh\to 4b$ process,
a multi-layer feed-forward artificial neural network (ANN),
also known as a {\it perceptron}.\footnote{This type of ANNs are the same
  as those used to parametrize Parton Distribution Functions
in the NNPDF global analyses~\cite{DelDebbio:2004qj,Ball:2008by,Ball:2011mu,Ball:2010de}.}
%
This family of ANNs is sometimes known as {\it deep} neural networks.
%
The basic idea is that the ANN is trained on an input composed on
all  signal and background
events which satisfy the requirements of the
cut-based analysis.
%
The main result of the trained ANN is identifying, in a fully automated way,
which are the most relevant variables to discriminate between the two.

In this work, the ANN that we use has the following architecture.
\be
\label{eq:nn1}
N_{\mathrm{var}}\times5\times3\times1 \, ,
\ee
where $N_{\mathrm{var}}$ represents the number of input variables for the MVA.
All neural-network layers use a sigmoid activation function, allowing
for a probabilistic
interpretation of the neural-network output, including in the final layer.
%
In Fig.~\ref{fig:nnarch} we show an example of the ANN that we use in this work, corresponding here
to the case of the resolved category.

@@

 Therefore, the training of the neural networks consist on the
 minimization of the cross-entropy error,
 Eq.~(\ref{cross-entropy}), which in this work is achieved using a
 Genetic Algorithm.
 %
 Since given the large number of events the probability of
 over-fitting is tiny, we just train the MVA for a very long
 number of generations, in this case $N_{\rm gen}=50k$.
 %
 We have verified that if a larger number of generations
 are used the results are unchanged.
 %
 We have also verified that using a cross-validation stopping
 criterion (in particular, the same one as
 that used in the NNPDF3.0 fit~\cite{Ball:2014uwa})
 does not modify the results.

@@

 To identify the most discriminatory variables, we have to train
 the MVA using a large number of kinematic distributions for
 signal and background events.
%
In our $hh\to 4b$ case,
the input variables for the MVA differ between our three categories,
and in particular in the case of large-$R$ jets we want to exploit
the richness of substructure information.

For the three categories, boosted, intermediate and resolved,
the following common variables are used as input to the MVA:
\begin{itemize}
\item The transverse momenta of the leading and subleading Higgs, $p_{T,h_1}$ and $p_{T,h_2}$.
\item The transverse momentum of the reconstructed Higgs pair, $p_{T,hh}$.
\item The invariant masses of the leading and sub-leading Higgs candidates, $m_{h,1}$ and $m_{h,2}$.
\item The invariant mass of the reconstructed Higgs pair, $m_{hh}$.
\item The separation in $R$ between the two Higgs candidates, $\Delta R_{hh}$.
\item The separation in $\phi$  between the two Higgs candidates, $\Delta \phi_{hh}$.
\item The separation in $\eta$  between the two Higgs candidates, $\Delta \eta_{hh}$.
\end{itemize}
In addition, in the boosted category we use
  the transverse momenta of the leading, $p_{T,h_{1,1}}$ and $p_{T,h_{1,2}}$ and
  sub-leading, $p_{T,h_{2,1}}$ and $p_{T,h_{2,1}}$, Higgs candidate subjets.
  %
  In the resolved category instead,
  the corresponding variables are
  the transverse momenta $p_{T,i}$ of the four leading 
  $b$-tagged small-$R$ jets in the event.
  %
  In the intermediate category, we use the
  transverse momenta of the subjets
  from the large-$R$ jet $p_{T,h_{1,1}}$ and $p_{T,h_{1,2}}$ and the
 transverse momenta $p_{T,i}$ of the two leading 
  $b$-tagged small-$R$ jets.
  %
Therefore, we have a total of 13 variables which are common to the three categories.
%

@@

  The ratio of energy correlation functions $C_2^{\beta}$ is defined as
  \be
  \label{eq:c2}
C_2^{(\beta)} \equiv \frac{ {\rm ECF}(3,\beta) {\rm ECF}(1,\beta)}{\lc {\rm ECF}(2,\beta)\rc ^2} \, ,
\ee
while $D_2^{(\beta)}$ is instead defined as a double ratio of ECFs, that is
\be
e_3^{(\beta)}\equiv \frac{ {\rm ECF}(3,\beta)}{\lc {\rm ECF}(1,\beta)\rc^3} \, , \quad
  e_2^{(\beta)}\equiv \frac{ {\rm ECF}(2,\beta)}{\lc {\rm ECF}(1,\beta)\rc^2} \, , \quad
  \label{eq:d2}
D_2^{(\beta)} \equiv \frac{ e_3^{(\beta)})}{\lp e_2^{(\beta)} \rp^3} \, .
\ee
The energy correlation functions ${\rm ECF}(N,\beta)$ are defined
  in~\cite{Larkoski:2013eya} with the motivation that $(N+1)$-point correlators
  are sensitive to $N$-prong substructure.
  %
  The free parameter $\beta$ is set to a value of $\beta=2$,
  as recommended by the authors of Refs.~\cite{Larkoski:2013eya,Larkoski:2014gra}.
\end{itemize}
%
Therefore, there are in total $N_{\mathrm{var}}=13$ variables for the resolved analysis,
17 for the semi-resolved, and 21 for the boosted category.
%
Given that the MVA is able to identify the most discriminatory variables
and to ignore those that have smaller effect, it is advantageous to
include as many substructure variables as possible.
%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.65\textwidth]{plots/Boosted_disc_noPU.pdf}
\includegraphics[width=0.48\textwidth]{plots/Intermediate_disc_noPU.pdf}
\includegraphics[width=0.48\textwidth]{plots/Resolved_disc_noPU.pdf}
\caption{\small The distribution as a function of the ANN output
  for the signal and background MC events in the three categories:
  boosted (upper plot), intermediate (lower left plot) and
  resolved (lower right plot).
  %
  All distributions are normalized so that their integral
  adds up to one.
}
\label{fig:nnresponse}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

The post-MVA results for $S/\sqrt{B}$ and $S/B$ as a function of the cut
in the ANN output for each of the three categories are shown in
Fig.~\ref{fig:sb_mva}.
%
Note that the values of $S/\sqrt{B}$ and $S/B$
for $y_{\rm cut}=0$ correspond to the values at
the end of the cut-based analysis, see
the {\bf C2} row of Table~\ref{tab:cutflow_noPU_1}.
%
Indeed, at the end of the cut based analysis $S/\sqrt{B}$ was
approximately
1.0, 0.5 and 0.5 for the boosted, intermediate and resolved
categories, as can be seen from Fig.~\ref{fig:sb_mva}.
%
The same applies to $S/B$.

@@

The results of Fig.~\ref{fig:sb_mva} are interesting in that they show
that for the resolved and boosted categories the improvement in signal
significance arising from the MVA is rather moderate, less that a factor two
in the two cases.
%
Even after the MVA, $S/\sqrt{B}$ never grows above 1.0.
%
The real power of the MVA can be seen for the boosted category,
where the signal significance rises from around 1.0 in the cut-based
analysis to above 3.0 for values of the ANN output cut $y_{\rm cut}\ge0.85$.
%
This value of the cut seems reasonable, since we know from Fig.~\ref{fig:nev2}
that the signal yield at the HL-LHC is still quite substantial,
of the order of 100 event.
%
Therefore, thanks to the MVA, we predict that one should be able
to find evidence for the production of Higgs boson pairs
at the HL-LHC using the $4b$ channel alone.

@@

Our final results are summarized in Table~\ref{table:cutflowMVA}, where we indicate
the value of the optimal ANN output
cut in each category,
the number of signal and background events $N_{\rm ev}$ expected
at the HL-LHC,
and also $S/\sqrt{B}$ and $S/B$.
%
For reference, we also include the results at the end of
the cut-based
analysis.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Boosted category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm cut}=0$} & \multicolumn{4}{c}{$y_{\rm cut}=0.86$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      193   &  $3.8\,10^4$    &  0.99    &  $5.1\,10^{-3}$  & 107 & 1040 & 3.31  & 0.10\\
        \hline
  \end{tabular}
   $\,$\\
  \vspace{0.4cm}
  \noindent
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Intermediate category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm cut}=0$} & \multicolumn{4}{c}{$y_{\rm cut}=0.49$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      660   &   $2.2\,10^6$   &   0.45   &  $3.1\,10^{-4}$ & 480  & $3.2\,10^5$ & 0.85  & $1.5\,10^{-3}$ \\
        \hline
  \end{tabular}
    $\,$\\
  \vspace{0.4cm}
  \noindent
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Resolved category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm cut}=0$} & \multicolumn{4}{c}{$y_{\rm cut}=0.17$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
    $1.6\,10^{3}$   &   $1.1\,10^{7}$   & 0.50     &  $1.5\,10^{-4}$   & 968 &
    $1.2\,10^{6}$  & 0.90 & $8.2\,10^{-4}$ \\
        \hline
  \end{tabular}
  \caption{\small The results for the number of signal and
    background events
    at the HL-LHC for both the cut-based analysis and after the
    optimal MVA cut, as well as the $S/\sqrt{B}$ and $S/B$
    values in the two cases.
    %
    From top to bottom we show the results for boosted,
    intermediate and resolved categories.
    %
    The value of the ANN output cut $y_{\rm cut}$
    is optimised separately in the
    three categories.
    \label{table:cutflowMVA}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

From Table~\ref{fig:sb_mva} it is also possible to
extract the signal significance that would be obtained
at the end of Run II with $\mathcal{L}_{\rm int}=300$ fb$^{-1}$:
in the boosted category we expect 11 signal events and 100 background
events, leading to a $S\sqrt{B}=1.1$, too small specially accounting
for realistic systematic experimental uncertainties.
%
So the full luminosity of the HL-LHC is really needed to exploit
the information contained in this final state, unless of course
new BSM dynamics enhance the di-Higgs production rates.

The results of Table~\ref{table:cutflowMVA} include
only the statistical uncertainty in the determination of
the background number of events.
%
We  briefly comment now how these results would be affected
if we also take into account the systematic experimental
uncertainty in the measurement
of the multijet backgrounds distributions.
%
While there is no unique definition of how signal significance is modified
by experimental systematic uncertainties~\cite{Bityukov:2002eq}, here we estimate
this effect with the simple correction
\be
\frac{S}{\sqrt{B}} \, \to \, \frac{S}{\sqrt{B}} \frac{1}{\sqrt{1+\lp \sigma^{\rm sys}\rp^2\cdot B}} \, ,
\ee
which corresponds to adding in quadrature to the statistical uncertainty  a single
uncorrelated systematic uncertainty $\sigma^{\rm sys}$.
%
Assuming an aggressive value of $\sigma^{\rm sys}=2\%$,
we find that the combined signal significance would be
reduced from 3.51 to 3.03.
%
While this quick estimate emphasizes the critical
importance of an experimentally accurate
measurement of multijet backgrounds at the LHC to fully exploit the $4b$
channel,
it is beyond the scope of this work to estimate realistically
the systematic uncertainties on QCD multijet distributions (and their bin-by-bin correlation)
that will be achieved at HL-LHC.

In the next section we discuss how the expectations for the HL-LHC
summarized in  Table~\ref{fig:sb_mva} are modified in different
conservative and aggressive scenarios for the performance of the
ATLAS and CMS experiments, in particular how the results depend
on the $b$-tagging performance and in the detector momentum
resolution.
%
But before that let us illustrate how the improvement in $S/\sqrt{B}$
obtained from the MVA in the boosted
category arises from the information contained in the
jet substructure variables.

>>>>

\section{Multivariate analysis}
\label{sec:mva}

From the cut-based analysis,
combining the significance of the three categories,
we obtain a signal significance of $S/\sqrt{B}\simeq 1.7 (3.5)$
with all backgrounds (only QCD $4b$) taken into account.
%
This is clearly
to be enough to be able to claim observation
of Higgs pair production
in this channel, and even less to extract the trilinear Higgs  coupling
$\lambda$, in particular
taking into account the that $S/B$ is at the permille level.
%
In this section we now show how this signal significance for
 can
be significance improved when
applying a multivariate analysis (MVA) subsequently
to the cut-based analysis.

%
Multivariate techniques are by now a mature tool to enhance signal
significance in HEP analysis, allowing
to open a new window to improve the performance
of many physics analysis and searches.
%
In particular, the classification of events as arising from either signal or
background processes via the use of a multivariate analysis is a
commonly applied strategy in LHC
applications~\cite{Baldi:2014pta,Aaltonen:2012qt,
  Wardrope:2014kya,Chatrchyan:2013zna,Dall'Osso:2015aia}.
%
For instance, the CMS $Vh(\to b\bar{b})$ analysis~\cite{Chatrchyan:2013zna}
is based on a Boosted Decision Tree (BDT) MVA, and the UCL $hh\to 4b$
feasibility study~\cite{Wardrope:2014kya}
also used  a BDT to improve signal discrimination.

In this section, first of all we present the specific MVA that we use,
based on feed-forward multi-layer neural networks.
%
We then introduce the input variables that are
used the MVA, including the jet substructure
variables, and then present the results of $S/\sqrt{B}$ including
the MVA effects.
%
Since we find that jet substructure variables are a crucial input
to the MVA, we finally show various kinematic distributions
of these variables.

@@

%
The specific type of  MVA that we use here, to
disentangle signal and background events in the $hh\to 4b$ process, will be
a multi-layer feed-forward artificial neural network (ANN),
also known as a {\it perceptron}.\footnote{This type of ANNs are the same
  as those used to parametrize Parton Distribution Functions
in the NNPDF global analyses~\cite{DelDebbio:2004qj,Ball:2008by,Ball:2011mu,Ball:2010de}.}
%
This family of ANNs is sometimes known as {\it deep} neural networks, since they have
a richer architecture with hidden layers.
%
The basic idea is that the ANN is trained on an input composed on
all  signal and background
events which satisfy the requirements of the
cut-based analysis.
%
The main result of the trained ANN is identifying, in a fully automated way,
which are the most relevant variables to discriminate between the two classes
of events.

In this work, the ANN that we use has the following architecture.
\be
\label{eq:nn1}
N_{\mathrm{var}}\times5\times3\times1 \, ,
\ee
where $N_{\mathrm{var}}$ represents the number of input variables for the MVA,
which is specific of the analysis category: resolved, intermediate, or boosted.
%
All neural-network layers use a sigmoid activation function, allowing
for a probabilistic
interpretation of the neural-network output, including in the final layer.
%
In Fig.~\ref{fig:nnarch} we show an example of the ANN that we use in this work, corresponding here
to the case of the boosted category (thus $N_{\mathrm{var}}=21$, as we explain below).

@@

 Therefore, the training of the neural networks consist on the
 minimization of the cross-entropy error,
 Eq.~(\ref{cross-entropy}), which in this work is achieved using a
 Genetic Algorithm.
 %
 The GA training is performed for a very large 
 number of generations, in this case $N_{\rm gen}=50k$, to avoid
 under-training.
 %
 We have verified that if a larger number of generations
 are used the results are unchanged.
 %
 We have also verified that using a cross-validation stopping
 criterion, in particular, the same one as
 that used in the NNPDF3.0 analysis~\cite{Ball:2014uwa}
 does not modify the results.
 %
 This cross-validation entails dividing the input MC dataset into two disjoint sets,
 and use one for the training and the other for the validation: the optimal
 stopping point is then indicated by the minimum of the error function
 Eq.~(\ref{cross-entropy}) for the validation sub-sample.

@@

 To identify the variables with higher discriminatory power,
 we need to train
 the MVA using a large number of kinematic distributions for
 signal and background events.
%
In our $hh\to 4b$ case,
the input variables for the MVA differ between our three categories,
and in particular in the case of large-$R$ jets (for the boosted
and intermediate categories) we want to exploit
the richness of substructure information.

For the three categories, boosted, intermediate and resolved,
the following common variables are used as input to the MVA:
\begin{itemize}
\item The transverse momenta of the leading and subleading Higgs, $p_{T,h_1}$ and $p_{T,h_2}$.
\item The transverse momentum of the reconstructed Higgs pair, $p_{T,hh}$.
\item The invariant masses of the leading and sub-leading Higgs candidates, $m_{h,1}$ and $m_{h,2}$.
\item The invariant mass of the reconstructed Higgs pair, $m_{hh}$.
\item The separation in the $\phi$--$\eta$ plane
  between the two Higgs candidates, $\Delta R_{hh}$.
  \item The separation in $\eta$  between the two Higgs candidates, $\Delta \eta_{hh}$.
\item The separation in $\phi$  between the two Higgs candidates, $\Delta \phi_{hh}$.
\end{itemize}
In addition, in the boosted category we use
  the transverse momenta of the leading, $p_{T,h_{1,1}}$ and $p_{T,h_{1,2}}$ and
  sub-leading, $p_{T,h_{2,1}}$ and $p_{T,h_{2,1}}$, Higgs candidate subjets.
  %
  In the resolved category instead,
  the corresponding variables are
  the transverse momenta $p_{T,i}$ of the four leading 
  $b$-tagged small-$R$ jets in the event.
  %
  In the intermediate category, we use the
  transverse momenta of the subjets
  from the large-$R$ jet $p_{T,h_{1,1}}$ and $p_{T,h_{1,2}}$ and the
 transverse momenta $p_{T,i}$ of the two leading 
  $b$-tagged small-$R$ jets.
  %
Therefore, we have a total of 13 variables which are common to the three categories.
%

@@

  The ratio of energy correlation functions $C_2^{\beta}$ is defined as
  \be
  \label{eq:c2}
C_2^{(\beta)} \equiv \frac{ {\rm ECF}(3,\beta) {\rm ECF}(1,\beta)}{\lc {\rm ECF}(2,\beta)\rc ^2} \, ,
\ee
while $D_2^{(\beta)}$ is instead defined as a double ratio of ECFs, that is
\be
e_3^{(\beta)}\equiv \frac{ {\rm ECF}(3,\beta)}{\lc {\rm ECF}(1,\beta)\rc^3} \, , \quad
  e_2^{(\beta)}\equiv \frac{ {\rm ECF}(2,\beta)}{\lc {\rm ECF}(1,\beta)\rc^2} \, , \quad
  \label{eq:d2}
D_2^{(\beta)} \equiv \frac{ e_3^{(\beta)})}{\lp e_2^{(\beta)} \rp^3} \, .
\ee
The energy correlation functions ${\rm ECF}(N,\beta)$ are defined
  in~\cite{Larkoski:2013eya} with the motivation that $(N+1)$-point correlators
  are sensitive to $N$-prong substructure.
  %
  The free parameter $\beta$ is set to a value of $\beta=2$,
  as recommended by the authors of Refs.~\cite{Larkoski:2013eya,Larkoski:2014gra}.
\end{itemize}
%
Therefore, there are in total $N_{\mathrm{var}}=13$ variables for the resolved analysis,
17 for the semi-resolved, and 21 for the boosted category.
%
Given that the MVA is able to identify the most discriminatory variables
and to ignore those that have smaller effect, it is advantageous to
include as many substructure variables as possible.
%
This is one of the main advantages of ANNs in this context: they are inherently redundant, so
adding additional information, even if carries very little weight, will never degrade
the classification power of the MVA.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.65\textwidth]{plots/Boosted_disc_noPU.pdf}
\includegraphics[width=0.48\textwidth]{plots/Intermediate_disc_noPU.pdf}
\includegraphics[width=0.48\textwidth]{plots/Resolved_disc_noPU.pdf}
\caption{\small The distributions, at the end of the
  GA training, 
  for the signal and background MC events in the three categories:
  boosted (upper plot), intermediate (lower left plot) and
  resolved (lower right plot), as a function of the ANN output.
  %
  All distributions are normalized so that their integrals
  add up to one.
}
\label{fig:nnresponse}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

These results are summarized in Table~\ref{table:cutflowMVA}, where we indicate
the value of the optimal ANN output
cut in each category,
the number of signal and background events $N_{\rm ev}$ expected
at the HL-LHC,
and also $S/\sqrt{B}$ and $S/B$.
%
For reference, we also include the results at the end of
the cut-based
analysis.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \small
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Boosted category}\\
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm cut}=0$} & \multicolumn{4}{c}{$y_{\rm cut}=0.82$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      1070   &  $7.6\cdot 10^5$    &  1.2    &  $1.4\cdot 10^{-3}$  & 790 & $2.2\cdot 10^4$ & 5.4  & 0.034\\
      \hline
      \multicolumn{8}{c}{$\quad$}\\
    \hline
    \multicolumn{8}{c}{Intermediate category}\\
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm cut}=0$} & \multicolumn{4}{c}{$y_{\rm cut}=0.80$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      670   &   $5.3\cdot 10^5$   &   0.9   &  $1.2\cdot 10^{-3}$ & 360  & $5.5\cdot 10^3$ & 4.8  & $0.06$ \\
      \hline
      \multicolumn{8}{c}{$\quad$}\\
    \hline
    \multicolumn{8}{c}{Resolved category}\\
    \hline
      \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm cut}=0$} & \multicolumn{4}{c}{$y_{\rm cut}=0.65$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
    $3700$   &   $1.5\cdot 10^{7}$   & 1.0     &  $3\cdot 10^{-4}$   & $1.3\cdot 10^{3}$ &
    $8.3\cdot 10^{4}$  & 4.5 & 0.01 \\
        \hline
  \end{tabular}
  \caption{\small Post-MVA results, for the optimal value of the
    ANN discriminant, in the three categories, compared with the
    result of the cut-based analysis.
    %
    We indicate the number of signal and
    background events
    at the HL-LHC, as well as the signal significance $S/\sqrt{B}$ and
    the signal over background ratio $S/B$.
    %
    The value of the ANN output cut $y_{\rm cut}$
    is optimised separately in the
    three categories.
    %
    The $y_{\rm cut}=0$ value corresponds to row C2 in Table~\ref{tab:cutflow_noPU_1}.
    \label{table:cutflowMVA}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The post-MVA results for $S/\sqrt{B}$ and $S/B$ as a function of the cut
in the ANN output for each of the three categories are shown in
Fig.~\ref{fig:sb_mva}.
%
Note that the values of $S/\sqrt{B}$ and $S/B$
for $y_{\rm cut}=0$ correspond to the values at
the end of the cut-based analysis, see
the C2 row of Table~\ref{tab:cutflow_noPU_1}.
%
Indeed, at the end of the cut based analysis, the
signal significance $S/\sqrt{B}$ was
1.2, 0.9 and 1.0 for the boosted, intermediate and resolved
categories respectively, consistent with Fig.~\ref{fig:sb_mva}.
%
The same applies to the signal over background ratio $S/B$.

@@

The results of Fig.~\ref{fig:sb_mva}  show that
for all categories the  improvement in signal
significance arising from the MVA is substantial, growing from around 1
to around 5 in the three cases.
%
The optimal values of $y_{\rm cut}$ ensure that the
number of events at the HL-LHC is still well appreciable: we obtain
1300, 360 and 800 events in the resolved, intermediate, and
boosted categories respectively.
%
Therefore, thanks to the MVA, we predict that one should be able
to find evidence for the production of Higgs boson pairs
at the HL-LHC using the $4b$ channel alone.

@@

@@

These remarkable results are however subject to an obvious criticism:
at the HL-LHC the intense PU is know to substantially affect
kinematical distributions, and in particular, it is expected to degrade
in some extent the results of the present analysis.
%
To preempt this criticism, in the next section
we discuss how the expectations for the HL-LHC
summarized in  Table~\ref{fig:sb_mva} are modified in the presence
of PU expected at the HL-LHC.
%
But before that let us illustrate how the improvement in $S/\sqrt{B}$
obtained from the MVA in the boosted
category arises from the information contained in the
jet substructure variables.
